### 2021-04-29 4-Technology.tex

precise  description  of  the  technology  / approach that is used

intro Openflexo ; intro model federation

jc