### 2021-05-07 1-Main.tex
Updates from Overleaf

\documentclass[american]{emisa}

\usepackage{todonotes}

\newcommand{\noteSM}[1]{\todo[color=red!40, author=\textbf{Salvador}, inline, caption={}]{#1}}
\newcommand{\noteJC}[1]{\todo[color=pink!40, author=\textbf{JC}, inline, caption={}]{#1}}
\newcommand{\noteSylvain}[1]{\todo[color=green!40, author=\textbf{Sylvain}, inline, caption={}]{#1}}
\newcommand{\noteAntoine}[1]{\todo[color=yellow!40, author=\textbf{Antoine}, inline, caption={}]{#1}}
\newcommand{\noteFabien}[1]{\todo[color=orange!40, author=\textbf{Fabien}, inline, caption={}]{#1}}
\newcommand{\noteJoel}[1]{\todo[color=blue!20, author=\textbf{Joel}, inline, caption={}]{#1}}

\newcommand{\mpc}{MULTI process challenge\xspace}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{tikz-uml}

%\usepackage{amsmath,mathtools}%amssymb,
\usepackage{algorithmic,algorithm}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\begin{document}
\begin{article}{
%\title{\mpc: the FML solution}
\title{Multi-level modelling with Openflexo/FML - A contribution to the \mpc}

%% Group authors per affiliation:
%\author{Sylvain Guérin, Joel Champeau}
%\address{ENSTA Bretagne}

%\author{Jean-Christophe Bach, Salvador Martinez, Antoine Beugnard, Fabien Dagnat}
%\address{IMT Atlantique}

%% or include affiliations in footnotes:
\author*{Sylvain Guérin}{firstname.lastname@ensta-bretagne.fr}
\address{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author{Joel Champeau}
\address[a]{}

\author{Jean-Christophe Bach}
\address{IMT Atlantique, Lab-STICC, UMR 6285, F-29238 Brest, France}
%\ead{firstname.lastname@imt-atlantique.fr}

\author{Salvador Mart\'inez}
\address[b]{}

\author{Antoine Beugnard}
\address[b]{}

\author{Fabien Dagnat}
\address[b]{}

\abstract{\input{2-Abstract}}
\keywords{Multilevel Modeling Challenge, Federation Modeling Language, OpenFlexo}
\bibliography{biblio.bib}

}
%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.

\section{Introduction}
\label{sec:introduction}
\input{3-Introduction.tex}

\noteSM{test}
\noteJC{test}
\noteSylvain{test}
\noteAntoine{test}
\noteFabien{test}
\noteJoel{test}

\section{Technology}
\label{sec:technology}
\input{4-Technology.tex}

\section{Analysis}
\label{sec:analysis}
\input{5-Analysis.tex}

\section{Model presentation}
\label{sec:model}
\input{6-ModelPresentation.tex}

\section{Satisfaction of Requirements}
\label{sec:requirements}
\input{7-RequirementsSatisfaction.tex}

\section{Assessment of the Modeling Solution}
\label{sec:discussion}
\input{8-Discussion.tex}

\section{Related work}
\label{sec:relatedwork}
\input{9-Relatedworks.tex}

\section{Conclusions}
\input{10-Conclusions.tex}

\printbibliography

\end{article}
\end{document}
%%