### 2021-05-19 7-RequirementsSatisfaction.tex
Updates from Overleaf

Now we discuss how the solution satisfies the challenge requirements.

Sylvain