### 2021-05-21 8-Discussion.tex

\noteJC{notes en vrac
  \begin{itemize}
    \item qui ? jc, fabien
    \item l'appel contraint plus ou monis la structure et le contenu de cette
      section
  \begin{itemize}
    \item discussing choices  made (ex : héritage vs attr), pointing out potential compromises/deficiencies ;
    \item \emph{Challenge respondents must discuss their multilevel
model solution with regard to the following aspects, each
of which should be treated in a specific sub-section of the
“Assessment” section of the article}
  \end{itemize}
    \item pas d'explicitation de niveaux de « concrétude » : conséquences ? facile de tracer des liens entre les niveaux ; difficile d'avoir le niveau ;
    \item outil adaptable
    \item niveau d'abstraction
    \item réutilisabilité des modèles ; 
    \item ajouter quelques métriques "gros grains"
    \item difficultés/facilités : contraintes faites à la main par rapport aux
      langages où on peut spécifier le niveau d'extension lorsque l'on trace un
      lien multi niveaux -> outil plus souple, mais moins de comportements
      automatisés
    \item + ouvert
    \item moins de support (faudrait implémenter ce qui manque)
    \item requiert de la méthode, ajouter des comportements, etc.
    \item point saillant :  outillage "gratuit" = on fournit une solution + l'outil pour exécuter le processus
\end{itemize}}

\todo[inline]{on pourrait aussi répondre ici aux 3 questions posées
explicitement (copiées dansl'intro pour le moment), a priori non, le chef
trouve que c'est chiant}

%
%%niveaux de ``concrétude'' -> level-blind solution
%
%Although our tool was not designed with a focus on multi-level modelling, we were able to propose a fully working solution. 
%
%Our approach is level-blind, therefore it is not easy to quickly retrieve the abstraction level of a concept.
%
%%métriques gros grains ?
%
%
%%choices, compromises/deficiences
%During the modeling phase, we had to choose 
%(for instance requirement xx 
%
%%+/-
%
%%points saillants

  %plan plus ou moins imposé par l'appel
\todo[inline]{mandatory discussion aspects (on verra plus tard si on fait des
sous-sections ou pas) ↓}

  \subsection{Basic modeling constructs}
  %Explain the basic modeling constructs used in the solution.

  \subsection{Levels}
  %Levels (or other model content organization schemes employed)

  %Explain the nature of “levels” in the model, how model elements are arranged
  %on these levels and which relationships (such as “instance-of”) may feature
  %between elements at different levels. The nature of levels should be
  %captured by explicitly stating the level segregation and the level cohesion
  %principles used [5]. Avoid vague language such as “higher level concepts are
  %more abstract” if the inter-level relationship is more specific. If the
  %inter-level relationship is deliberately allowed to be vague, state this
  %explicitly.

  The Openflexo approach is level-blind: ``levels'' have no specific
  nature and there are no numbered levels.

  \subsection{Number of levels}
  %Elaborate whether the submitted solution could have had more or fewer levels
  %and explain how any potentially existing degrees of freedom were resolved.

  Due to the fact our approach is level-blind, our solution could have had more
  or fewer levels depending on the variations of the use case. %As levels are
  %not specifically encoded,

  \subsection{Cross-level relashionships}

  %Discuss if and how associations and links can connect model elements at
  %different levels. State well-formedness constraints, if any apply.

  Thanks to the level-blindness nature of the Openflexo approach, cross-level
  relashionships are not an issue. Model elements of different levels can be
  linked each other in a transparent way. %If the user has defined
  %well-formedness constraints, then they are

  \subsection{Cross-level constraints}

%Discuss if and how constraints can span multiple levels, especially with
%regard to cross-level relationships.

  As for cross-level relationships, cross-level constraints do not have a
  specific nature. They are like any constraint a user can define on a model
  element. Therefore, if a constraint is mandatory when establishing a
  relationship between elements from different levels, the user has create it
  manually. This can be a limit of our approach: the flexibility of our tooling
  at the cost of a limited number fo automated behaviors.

  \subsection{Integrity mechanism}

  %Discuss how the integrity of level contents is preserved when changes to
  %level contents occur.

  In order to preserve the integrity of level contents when changes to level
  contents occur, a validation has to be triggered. The tool reports all
  inconsistencies that the user is free to fix.

  \subsection{Deep characterization}

  %Discuss if and how higher levels influence elements at lower levels with a
  %level distance of two or more. Such an influence may be desired to ensure
  %properties of lower level elements regardless of the design choices that
  %modelers make at intermediate levels, including future extensions to
  %intermediate levels.

  \noteJC{pas trop clair pour moi pour le moment}

  \subsection{Generality}

  %Discuss the generality of the solution. Is (part of) it applicable to other
  %domains? Does it embody invariant principles of the domain(s) it covers with
  %minimal redundancy?

  Due to the nature of model federation, the models are highly reusable. Our
  solution separates the domain-specific elements from general purpose ones,
  making easily possible to apply it to other domains. 
  \noteJC{à développer + ajouter avantage de l'outil gratuit ici ?}

  \subsection{Extensibility}

  %Elaborate how the solution would respond to further requirements, such as
  %further special tasks that must be taken care of by special actors. Identify
  %expected extension points in the solution, e.g., subtyping opportunities. If
  %level insertion is a possibility in your chosen approach, explain how it
  %would be performed.

  A strength of the approach of model federation we have adopted resides in its
  flexibility. One can extend a solution easily:

  \noteJC{comment étendre : dépend un peu des extensions, ne prendre que celle
  donnée en exemple ? ; level insertion possible : oui, vu que pas de notion de
level}

\todo[inline]{recommended discussion aspects ↓}
\subsection{formalisms and tools ?}

%Indicate whether there are formalisms to establish the semantics of the MLM
%technique and/or tools that support the present solution

\noteJC{on a du tooling, un peu moins de formalisme}

\subsection{model verification}

%Discuss model verification (e.g., consistency analyses) or other quality
%assessment mechanisms supported by the MLM technique employed.