### 2021-05-25 Figures/federation.tex
model federation figure

\documentclass{standalone}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{graphicx}

\usepackage{tikz}
\usetikzlibrary{calc,shapes.geometric,arrows.meta,backgrounds}

\begin{document}
  \begin{tikzpicture}[ >=Classical TikZ Rightarrow]
    \draw[use as bounding box] (-4.065,-4.065) rectangle (4.065,4.065);
    \foreach \n/\h in {0/1.2cm,1/1.3cm,2/1.2cm,3/1.1cm,4/1cm}{
      \node[inner sep=0pt] (m\n) at (90+\n*360/5: 3cm) {\includegraphics[height=\h]{house\n}};
      \node[minimum width=3mm,minimum height=2mm,draw=orange,fill=orange!30,inner sep=0pt,rotate=\n*360/5] (ms\n) at (90+\n*360/5:1.18cm) {};
      \draw[<->] (m\n) -- (ms\n);
    }
    \begin{pgfonlayer}{background}
      \foreach \n/\c in {0/black!15,1/black!20,2/black!25,3/black!20,4/black!25}{
        \node[inner sep=1pt,outer sep=0pt,circular sector,circular sector angle=360/5,fill=\c,minimum width=6.8cm,rotate=-90+\n*360/5] (ts\n) at (90+\n*360/5: 2.73cm) {};
      }
      \begin{scope}[blue]
        \node[circle,draw=orange,fill=orange!30,text width=2cm,align=center,inner sep=1mm] {conceptual model};
        \node[anchor=center,text width=2cm,align=center] (ts) at (m0-|m4)  {technological spaces};
        \node[align=center,text width=0.92cm] (ms) at ($(m3)!0.5!(m4)$)  {model slots};
        \node[anchor=south east,text width=1.4cm,align=center] (fm) at ($(m0)!0.5!(m1)$)  {federated models};
        \draw[->,thin] (ts.west) -- +(-0.4,0);
        \draw[->,thin] (ts.south) -- +(0,-0.4);
        \draw[->,thin] (ms) -- (ms3);
        \draw[->,thin] (ms) -- (ms4);
        \draw[->,thin] (fm) -- (m0);
        \draw[->,thin] (fm) -- (m1);
      \end{scope}
    \end{pgfonlayer}
  \end{tikzpicture}
\end{document}