### 2021-05-31 1-main.tex
fix keywords command

\documentclass[american]{emisa}

\usepackage{todonotes}

\newcommand{\noteSM}[1]{\todo[color=red!40, author=\textbf{Salvador}, inline, caption={}]{#1}}
\newcommand{\noteJC}[1]{\todo[color=pink!40, author=\textbf{JC}, inline, caption={}]{#1}}
\newcommand{\noteSylvain}[1]{\todo[color=green!40, author=\textbf{Sylvain}, inline, caption={}]{#1}}
\newcommand{\noteAntoine}[1]{\todo[color=yellow!40, author=\textbf{Antoine}, inline, caption={}]{#1}}
\newcommand{\noteFabien}[1]{\todo[color=orange!40, author=\textbf{Fabien}, inline, caption={}]{#1}}
\newcommand{\noteJoel}[1]{\todo[color=blue!20, author=\textbf{Joel}, inline, caption={}]{#1}}

\emisainitialism{\EMF}{EMF}
\emisainitialism{\MOF}{MOF}
\emisainitialism{\MDE}{MDE}
\emisainitialism{\FML}{FML}

\newcommand{\mpc}{MULTI process challenge\xspace}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{tikz-uml}

%\usepackage{amsmath,mathtools}%amssymb,
\usepackage{algorithmic,algorithm}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\DeclareTextFontCommand{\mytexttt}{\ttfamily\hyphenchar\font=45\relax}

\lstdefinelanguage{fml}{
  keywords={
    abstract,boolean,break,byte,case,catch,char,class,%
    const,continue,default,do,double,else,extends,false,final,%
    finally,float,for,goto,if,implements,import,instanceof,int,%
    interface,label,long,native,new,null,package,private,protected,%
    public,return,short,static,super,switch,this,throw,%
    throws,transient,true,try,void,volatile,while,String},%
  morekeywords={
    [3]concept,model,forEach,assert,%
  },%
  sensitive,%
  morecomment=[l]//,%
  morecomment=[s]{/*}{*/},%
  morestring=[b]",%
  morestring=[b]',%
}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\scriptsize,
    breakatwhitespace=false,
    breaklines=true,
    flexiblecolumns=true,
    captionpos=b,
    emptylines=0,
    tabsize=2
}

\lstset{style=mystyle,language=fml}

\begin{document}
\begin{article}{
    % \title{\mpc: the FML solution}
    \title{Multi-level modeling with Openflexo/\FML - A contribution to the \mpc}

    %% or include affiliations in footnotes:
    \author{Sylvain Guérin}%{firstname.lastname@ensta-bretagne.fr}
    \address{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}
    % \ead{firstname.lastname@ensta-bretagne.fr}

    \author{Joel Champeau}
    \address[a]{}

    \author*{Jean-Christophe Bach}{jc.bach@imt-atlantique.fr}
    \address{IMT Atlantique, Lab-STICC, UMR 6285, 29238 Brest, France}
    % \ead{firstname.lastname@imt-atlantique.fr}

    \author{Salvador Mart\'inez}
    \address[b]{}

    \author{Antoine Beugnard}
    \address[b]{}

    \author{Fabien Dagnat}
    \address[b]{}

    \abstract{\input{2-abstract}}
    \keywords{Multilevel Modeling Challenge \and Federation Modeling Language
    \and Openflexo}
    \bibliography{biblio.bib}
}

\section{Introduction}
\label{sec:introduction}
\input{3-introduction.tex}

%\noteSM{test}
%\noteJC{test}
%\noteSylvain{test}
%\noteAntoine{test}
%\noteFabien{test}
%\noteJoel{test}

\section{Technology}
\label{sec:technology}
\input{4-technology.tex}

\section{Analysis}
\label{sec:analysis}
\input{5-analysis.tex}

\section{Model presentation}
\label{sec:model}
\input{6-modelpresentation.tex}

\section{Satisfaction of requirements}
\label{sec:requirements}
\input{7-requirementssatisfaction.tex}

\section{Assessment of the modeling solution}
\label{sec:discussion}
\input{8-discussion.tex}

\section{Related work}
\label{sec:relatedwork}
\input{9-relatedworks.tex}

\section{Conclusions}
\label{sec:conclusions}
\input{10-conclusions.tex}

\printbibliography

\end{article}

\end{document}
%%